---
title: "Sentiment Analysis on News Sources Regarding Electric Vehicles"
author: "Hamza Khan"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # default line for .rmd files
```

Libraries
```{r message=FALSE, warning=FALSE}
library(tidytext) # for sentiment analysis
library(tidyverse) # for using read_csv, pipe, filter, ggplot2 and more
```

Reading data file
```{r message=FALSE, warning=FALSE}
setwd("C:/Users/Hamza/Desktop/FA2023/LING 460/sentiment-analysis-ev") # setting the working directory
data <- read_csv("ling_460_project_data.csv") # reading the data file
data1 <- data[, c(-1, -5, -6)] # removing unnecessary columns
```

Unnesting tokens
```{r message=FALSE, warning=FALSE}
data_unnested <- data1 %>% # creating a data frame named data_unnested from data1
  unnest_tokens(word, text_article) %>%  # unnesting tokens
  anti_join(stop_words) # removing stop words
```

BING Sentiment and recoding positive as 1, negative as 0
```{r message=FALSE, warning=FALSE}
data_bing <- data_unnested %>% # creating a data frame named data_bing from data_unnested
  inner_join(get_sentiments("bing")) # using inner_join to match sentiments from the bing lexicon
data_bing$score <- ifelse(data_bing$sentiment == "positive", 1, 0) # creating a new variable in the data frame which is the score variable regrades
```

Summarizing data by source_article, political_lean
```{r message=FALSE, warning=FALSE}
data_bing_source <- data_bing %>% # creating a new data frame from data_bing
  group_by(source_article, political_lean) %>% # Grouping by source_article and bringing in political_lean variable
  summarize(avg = mean(score)) # calculating the average for each source_article

data_bing_source <- arrange(data_bing_source, desc(avg)) # sorting the data frame descending by average

data_bing_lean <- data_bing %>% # creating a new data frame from data_bing
  group_by(political_lean) %>% # Grouping by the political_lean variable
  summarize(avg = mean(score)) # calculating the average for each source_article

data_bing_lean <- arrange(data_bing_lean, desc(avg)) # sorting the data frame descending by average
```

Graphing data_bing_source and data_bing_lean
  This revealed that left news sources have higher sentiment than news sources without a lean, and then right leaning news sources. Although it varies for sources in the middle and right.
```{r message=FALSE, warning=FALSE}
# creating bar graph for data_bing_source with legend and sorted by avg
ggplot(data = data_bing_source, aes(reorder(source_article, avg), avg, fill = political_lean)) +
  geom_col() + # bar chart
  coord_flip() + # making it horizontal for legibility
  ggtitle("Average Sentiment in Article by News Source") + # title
  theme_minimal() + # theme
  ylab("Average Sentiment (BING)") + # x-axis label (coord_flip is why this is x-axis)
  xlab("News Source") + # y-axis label (coord_flip is why this is y-axis)
  scale_y_continuous(limits = c(0,1)) + # setting scales for x-axis (coord flipped)
  geom_hline(yintercept = .5, linetype='dotted', col = 'red') + # creating a vertical line at the .5 mark (coord flipped)
  scale_fill_discrete(name = "Political Lean") + # labeling legend
  scale_fill_manual(values=c("blue", "gray", "red")) # setting colors

# creating bar graph for data_bing_lean with legend and sorted by avg
ggplot(data = data_bing_lean, aes(reorder(political_lean, avg), avg, fill = political_lean)) +
  geom_col() + # bar chart
  coord_flip() + # making it horizontal for legibility
  ggtitle("Average Sentiment by Political Lean of News Source") + # title
  theme_minimal() + # theme
  ylab("Average Sentiment (BING)") + # x-axis label (coord_flip is why this is x-axis)
  xlab("Political Lean") + # y-axis label (coord_flip is why this is y-axis)
  scale_y_continuous(limits = c(0,1)) + # setting scales for x-axis (coord flipped)
  geom_hline(yintercept = .5, linetype='dotted', col = 'red') + # creating a vertical line at the .5 mark (coord flipped)
  scale_fill_discrete(name = "Political Lean") + # labeling legend
  scale_fill_manual(values=c("blue", "gray", "red")) # setting colors
```

Three separate dataframes, one for each political lean
```{r message=FALSE, warning=FALSE}
left <- filter(data_bing, political_lean == "left") # filtering data_bing where political_lean is left
middle <- filter(data_bing, political_lean == "middle") # filtering data_bing where political_lean is middle
right <- filter(data_bing, political_lean == "right") # filtering data_bing where political_lean is right
```

T-test (welch due to difference in sample sizes)
  - H0: There are no differences in sentiment between news sources of different political leans
  - HA: There are differences in sentiment between news sources of different political leans
  
  - Left and Right: lowest p-value indicating that different political leans have differences in sentiment
  - Left and Middle: low p-value indicating that different political leans have differences in sentiment
  - Middle and Right: high p-value (above 0.05) indicating that different political leans don't have differences in sentiment
```{r message=FALSE, warning=FALSE}
t.test(left$score, right$score) # calculating t-test for the left and right columns
t.test(left$score, middle$score) # calculating t-test for the left and middle columns
t.test(middle$score, right$score) # calculating t-test for the middle and right columns
```

Creating dummy variable in data_bing_source to use in making glm
  I made three dummy variables instead of two for the glm, otherwise I would have done two.
```{r message=FALSE, warning=FALSE}
data_bing_source$left <- ifelse(data_bing_source$political_lean == "left", 1, 0) # creating a column in the data_bing_source data frame named left that is one if the political_lean is left and 0 otherwise
data_bing_source$right <- ifelse(data_bing_source$political_lean == "right", 1, 0) # creating a column in the data_bing_source data frame named right that is one if the political_lean is right and 0 otherwise
data_bing_source$middle <- ifelse(data_bing_source$political_lean == "middle", 1, 0) # creating a column in the data_bing_source data frame named middle that is one if the political_lean is middle and 0 otherwise
```

Logistic regression
```{r message=FALSE, warning=FALSE}
summary(glm(left~avg, data = data_bing_source)) # low p-values # looking at the summary for the logistic regression between left and avg variables
summary(glm(right~avg, data = data_bing_source)) # not good # looking at the summary for the logistic regression between right and avg variables
summary(glm(middle~avg, data = data_bing_source)) # even worse # looking at the summary for the logistic regression between middle and avg variables
```

###################################################################
## ## Going to do the same thing again with the AFINN lexicon ## ##
###################################################################

AFINN sentiment
```{r message=FALSE, warning=FALSE}
data_afinn <- data_unnested %>%  # creating a data frame named data_afinn from data_unnested
  inner_join(get_sentiments("afinn")) # using inner_join to match sentiments from the afinn lexicon
```

Summarizing data from afinn sentiment
```{r message=FALSE, warning=FALSE}
data_afinn_source <- data_afinn %>% # creating a new data frame from data_afinn
  group_by(source_article, political_lean) %>% # Grouping by source_article and bringing in political_lean variable
  summarize(avg = mean(value)) # calculating the average for each source_article

data_afinn_source <- arrange(data_afinn_source, desc(avg)) # sorting the data frame descending by average

data_afinn_lean <- data_afinn %>% # creating a new data frame from data_afinn
  group_by(political_lean) %>% # Grouping by the political_lean variable
  summarize(avg = mean(value)) # calculating the average for each source_article

data_afinn_lean <- arrange(data_afinn_lean, desc(avg)) # sorting the data frame descending by average
```

Creating bar graphs
```{r message=FALSE, warning=FALSE}
# creating bar graph for data_afinn_source with legend and sorted by avg
ggplot(data = data_afinn_source, aes(reorder(source_article, avg), avg, fill = political_lean)) +
  geom_col() + # bar chart
  coord_flip() + # making it horizontal for legibility
  ggtitle("Average Sentiment in Article by News Source") + # title
  theme_minimal() + # theme
  ylab("Average Sentiment (AFINN)") + # x-axis label (coord_flip is why this is x-axis)
  xlab("News Source") + # y-axis label (coord_flip is why this is y-axis)
  scale_y_continuous(limits = c(-1,1)) + # setting scales for x-axis (coord flipped)
  scale_fill_discrete(name = "Political Lean") + # renaming legend
  scale_fill_manual(values=c("blue", "gray", "red")) # setting colors
# ggsave("pic1.jpg") # saving the output as jpg

# creating bar graph for data_afinn_lean with legend and sorted by avg
ggplot(data = data_afinn_lean, aes(reorder(political_lean, avg), avg, fill = political_lean)) +
  geom_col() + # bar chart
  coord_flip() + # making it horizontal for legibility
  ggtitle("Average Sentiment by Political Lean of News Source") + # title
  theme_minimal() + # theme
  ylab("Average Sentiment (AFINN)") + # x-axis label (coord_flip is why this is x-axis)
  xlab("Political Lean") + # y-axis label (coord_flip is why this is y-axis)
  scale_y_continuous(limits = c(-1,1)) + # setting scales for x-axis (coord flipped)
  scale_fill_manual(values=c("blue", "gray", "red")) + # setting colors
  theme(legend.position="none") # hiding legend
# ggsave("pic2.jpg") # saving the output as jpg
```

Creating new data frames for each lean with afinn lexicon
```{r message=FALSE, warning=FALSE}
left_afinn <- filter(data_afinn, political_lean == "left") # filtering data_afinn where political_lean is left
middle_afinn <- filter(data_afinn, political_lean == "middle") # filtering data_afinn where political_lean is left
right_afinn <- filter(data_afinn, political_lean == "right") # filtering data_afinn where political_lean is left
```

T-tests
```{r message=FALSE, warning=FALSE}
t.test(left_afinn$value, right_afinn$value) # lowest p-value  # calculating t-test for the left and right columns
t.test(left_afinn$value, middle_afinn$value) # low p-value  # calculating t-test for the left and middle columns
t.test(middle_afinn$value, right_afinn$value) # just above .05  # calculating t-test for the middle and right columns
```

Creating dummy variables for glm (making 3 instead of 2 because of glm)
```{r message=FALSE, warning=FALSE}
data_afinn_source$left <- ifelse(data_afinn_source$political_lean == "left", 1, 0) # creating a column in the data_afinn_source data frame named left that is one if the political_lean is left and 0 otherwise
data_afinn_source$right <- ifelse(data_afinn_source$political_lean == "right", 1, 0) # creating a column in the data_afinn_source data frame named right that is one if the political_lean is right and 0 otherwise
data_afinn_source$middle <- ifelse(data_afinn_source$political_lean == "middle", 1, 0) # creating a column in the data_afinn_source data frame named middle that is one if the political_lean is middle and 0 otherwise
```

Logistic regression
```{r message=FALSE, warning=FALSE}
mod1glm <- glm(left~avg, data = data_afinn_source)
summary(mod1glm) # low p-value # looking at the summary for the logistic regression between left and avg variables
summary(glm(right~avg, data = data_afinn_source)) # barely low p-value # looking at the summary for the logistic regression between right and avg variables
summary(glm(middle~avg, data = data_afinn_source)) # bad p-values # looking at the summary for the logistic regression between middle and avg variables
```

Plotting mod1glm
```{r message=FALSE, warning=FALSE}
# Creating a scatter plot with a curve for glmmod1
ggplot(data_afinn_source, aes(x = avg, y = left)) + 
  geom_point() + # scatter plot
  stat_smooth(method = "glm", color = "red", se = FALSE, method.args = list(family = binomial)) #curve
```

Predict our results back on our original data and check for accuracy
  - It predicted 5 left and 4 not left
  - The not lefts are predicted correctly to not be left
  - All the lefts are predicted to be left, however, 2 extra are predicted to be left when they aren't
```{r message=FALSE, warning=FALSE}
mod1glm_pred <- predict(mod1glm, data_afinn_source)
mod1glm_pred
data_afinn_source$avg
```

